The Current State of Test-First vs. Test-Last Testing Draft

### 2013-11-09 Paper_Draft.tex
Draft for friday night.

When it comes to software development perhaps one of the most important and time consuming processes is that of software testing. In fact, early studies on software testing estimated that software testing could consume fifty percent or more of development costs for a product.  Because of this, the ability to optimize testing to reduce testing costs can be very valuable.  In this paper  we compare two popular methods, test-last testing, often used in waterfall software development methods, and test-first testing, often used in Agile test driven development software development methods, by reviewing recent studies on the subject.  In this review we discuss the possible benefits of test-first and test-last testing and possible problems with the current data comparing these two testing methods.  After that, we explore other methods in test-first testing besides test driven development, such as behavior driven development, in an attempt to find a better test-first testing model.  In the end we discuss our results and potential future studies to help clarify current data.
\end{abstract}

@@

\section{Introduction}
When it comes to software development, perhaps one of the most important and time consuming processes is software testing.  In fact, some early studies on software testing estimated that it could consume fifty percent or more of development costs for a product ~\cite{Bertolino:2007}.  because of this, businesses over time have become increasingly interested in attempting to optimize testing to reduce development costs.

@@

Currently, there is much debate in the testing community about whether or not test first or test last testing is superior.  The goal of this paper is to attempt to give an overview of the current state of this debate by analyzing current research data concerning the advantages and disadvantages of each of these testing methods.  Because research often lags behind current implementations and the field of test-first testing is currently changing due to its relatively new implementation, this paper will also explore where new test-first methods are going and compare them to the research given.  This paper will therefore discuss the advantages and disadvantages of test first testing versus test last testing and explore new emerging test-first methodologies in an attempt to determine the applicability of each methodology.

The rest of the paper is divided into four sections which will discuss the specific ideas above in more detail. In section two of the paper we will discuss what software testing is and current software development models with their supporting testing methods: test-first and test-last testing.  In section three we will provide an analysis of the data explaining the potential advantages and disadvantages of test-first testing compared to test-last testing.  In section four we will go over new test-first based testing methodologies and in section five we will state our conclusions and suggestions for further research in the field.

@@

When it comes to judging the effectiveness of certain testing methods it is important to note that there is no one standard quantitative measurement.  Because of this, many different types of measurement are used in an attempt to prove that one testing methodology is better then another.  Although there are many different attributes that are used for arguing the superiority of one method over another, we will only focus on three attributes that are commonly found throughout research; code coverage, testing time, and code correctness.  Code coverage refers to how many lines of code percent-wise are currently being tested. For example, if I write a test with 70\% test coverage it means that 70\% of the lines of code I wrote are being tested.  Testing time simply means how long it takes to write the actual tests.  Code correctness refers to how many errors are found within code after it is considered finished.  These three attributes are popular to use because they can be quantitatively measured and are considered important within the testing community.

@@

From as early as the 1970s a popular software development model was the waterfall model where software was developed in a series of phases.  The waterfall method has been a popular development model because it is considered simple to implement correctly and is also considered to be time efficient because completing previous phases helps latter phases become more efficient. The first phase in the standard waterfall model was the requirements phase where requirements were set by the customer or design company.  Next the design phase occurred where the product was designed which was then followed by the implementation phase where the product was actually created.  The final phase of the development phase was then the verification phase where testing and debugging occurred \mycomment{wikipedia waterfall model}.  Due to the fact that these processes often put testing at the end of development, a certain type of testing method, the test-last method, became very popular to use with the waterfall development method.

Test last testing is currently a popular testing method and is usually the first testing method that people tend to implement.  Test Last testing is the practice of writing tests after code has been written to check the functionality of the written code.  These tests are then used to check the written code for any potential errors which are then fixed by the developer until no errors are found by the tests.  Test last testing is considered simple to implement and can allow the programmer to focus on writing tests based on areas of his code that he may be concerned about.

@@

In the late 1990s some software developers started to criticize the sequence-oriented waterfall model, complaining that it was too brittle and inflexible to meet the demands of the standard customer.  In response to these critiques of the waterfall model a new model for developing software emerged, the agile development model.  This new development model, based on the tenets of the agile manifesto \mycomment{cite manifesto here!}, promoted the idea that all actions of development should not occur in an ordered sequence but in a series of time-boxed iterations where in each iteration setting requirements, product design, product implementation and product testing would occur simultaneously.  The goal of each iteration is to produce a demonstrable working product to show a customer and receive feedback on the product.  Some current development practices that are considered agile are Extreme Programming and Scrum.

@@

The most well known and most used test-first model is that of test driven development or TDD for short.  In its current use in the field, the phrase TDD is mostly used as a blanket term for any sort of development practice that uses test-first testing to drive code development. That being said it is important to note that although TDD has been turned into a vague term in this paper we will focus more on its original specific definition.  \mycomment{add more once we get to section 3}

\section{Research Data on Test-First vs Test-Last Testing}
In this section we attempt to attempt to obtain useful conclusions between test-first and test-last testing by observing current research articles.  In order to accomplish this we will first start with a data subsection that will go over the three main studies that will be used to draw our conclusions.  Next we will have an analysis subsection that will summarize the results of the study in terms of our three main empirical measurements: code coverage, code correctness, and total development time.  We will then have a discussion subsection that will take about potential issues with summarizing the data given in the research articles and then we shall proceed with a conclusion subsection where we will talk about potential conclusions we can draw from the data.

\subsection{Data}
 In 2012 Lemos et. al ~\cite{Lemos:2012} conducted a study on sixth semester students to see if test-first testing would increase code coverage, increase testing code size, increase code correctness, and/or decrease in total development time in auxiliary functions (functions with 10-200 lines of code) compared to test-Last testing.  In this study 39 sixth semester students, knowledgeable in testing techniques and participants of two 100 minute long test first modules, were asked to complete coding challenges over two sessions.  In the first session, half the students implemented testing-first testing while the other half implemented test-last testing to help solve the problem.  In the second section the students were asked to switch roles and given a different coding challenge.  In order to reduce domain knowledge bias amongst the coding challenges three different challenge question types were handed out at each session.  The results of the study, which were analyzed for statistical significance using the Wilcoxon/Mann-Whitney non-parametric signed-rank paired test, found that test-first testing significantly increased code coverage, testing code size and code development time with no significant increase in code correctness.

In 2010 Kollanus ~\cite{Kollanus:2010} conducted a literary review of TDD.  Using standard literary review methods, Kollanus found and reviewed forty different experiments in scientific journals, magazines, and conference proceedings that provided empirical evidence for TDD.  In the review Kollanus found that the most used measurements in these experiments could be clumped into three measurements: external code quality, internal code quality, and productivity.  In the study Kollanus observed that there was weak support for improved external code quality, very weak support for internal quality, and support for an increased amount of development time.

In 2012 Hellmann et. al conducted a summary on the current status of research of TDD.  Within Hellmann’s background there is a mention of a summary study produced by Jeffries and Mitnik that found in general that TDD largely resulted in an increase in quality, but one study they identified showed instead that TDD resulted in a strong negative impact on quality.  Additionality while they showed that TDD could reduce the amount of effort required by up to 27\%, most studies found an increase in effort of up to 100\%.

@@

@@

\subsection{Test Driven Development}
Test driven development, as mentioned earlier is currently the most widely used test first testing and design methodology.  Although in the previous cases we have been using TDD as a vague term in this section we will focus on the the original TDD methodology introduced in Extreme Programming.  In the original TDD methodology the developer repeats a step of processes to develop his code.  The process begins once all current tests pass or the new code has just been started.  When this occurs the programmer writes a new failing test that tests the simplest functionality the programmer wishes to add to their code.  Once the test has been written and the test fails, the programmer then writes the minimal amount of code to make the test pass.  After this step the programmer proceeds to refactor any clutter within his solution and then repeats the process until the code is complete.

Although TDD is a very popular style of test first testing, it should be noted that a perceived weakness to the TDD methodology exists.  The main issue many people have with test driven development is that it fails to suggest a style of testing and how to best define an independent functionality in ones code plan.  In the study ~\cite{Kettunen:2010} (cite others here as needed) people noted that TDD was difficult because they had a hard time deciding how to implement their tests effectively.  Another related issue noted in ~\cite{Kettunen:2010} was that some individuals found it hard to figure out what needed to specifically be tested thus some found they wrote an overabundance of tests causing the process to become inefficient and thus take more time.  Others found that focusing on the simplest functions first left serious potential problems to the end of development causing a serious time crunch problem for product release.  Due to these issues with defining functionality and other issues within TDD new methods for implementing test first testing are being designed and implemented in an attempt to solve these problems.
\subsection{Acceptance Test Driven Development}

>>>>

When it comes to software development, perhaps one of the most important and time consuming processes is that of software testing. In fact, early studies on software testing estimated that it could consume fifty percent or more of development costs for a product.  Because of this, the ability to optimize testing to reduce testing costs can be very valuable.  In this paper  we compare two popular methods, test-last testing, often used in waterfall software development methods, and test-first testing, often used in Agile test driven software development methods, by reviewing recent studies on the subject.  In this review we discuss the possible benefits of test-first and test-last testing and possible problems with the current data comparing these two testing methods.  After that, we explore other methods in test-first testing besides test driven development, such as behavior driven development, in an attempt to find a better test-first testing model.  In the end we discuss our results and potential future studies to help clarify current data.
\end{abstract}

@@

\section{Introduction}
When it comes to software development, perhaps one of the most important and time consuming processes is software testing.  In fact, some early studies on software testing estimated that it could consume fifty percent or more of development costs for a product ~\cite{Bertolino:2007}.  Because of this, software developers have become increasingly interested in attempting to optimize testing to reduce development costs.

@@

Currently, there is much debate in the testing community about whether or not test-first or test-last testing is superior.  The goal of this paper is to attempt to give an overview of the current state of this debate by analyzing current research data concerning the advantages and disadvantages of each of these testing methods.  Because research often lags behind current implementations and the field of test-first testing is currently changing due to its relatively new implementation, this paper will also explore where new test-first methods are going and compare them to the research given.  This paper will therefore discuss the advantages and disadvantages of test first testing versus test last testing and explore new emerging test-first methodologies in an attempt to determine the applicability of each methodology.

The rest of the paper is divided into four sections.  In section two of the paper we discuss what software testing is and current software development models with their supporting testing methods: test-first and test-last testing.  In section three we will provide an analysis of the data explaining the potential advantages and disadvantages of test-first testing compared to test-last testing.  In section four we will go over new test-first based testing methodologies and in section five we will state our conclusions and suggestions for further research in the field.

@@

When it comes to judging the effectiveness of certain testing methods it is important to note that there is no one standard quantitative measurement.  Because of this, many different types of measurement are used in an attempt to prove that one testing methodology is better then another.  Although there are many different attributes that are used for arguing the superiority of one testing method over another, we will focus on three attributes that are commonly found throughout research; code coverage, total development time, and code correctness.  Code coverage refers to how many lines of code percent-wise are currently being tested. For example, if I write a test with 70\% test coverage it means that 70\% of the lines of code that were wrote are being tested.  Total development time refers to how long it takes to finish a development project.  Code correctness refers to how many errors are found within code after it is considered finished.  These three attributes are popular to use because they can be quantitatively measured and are considered important within the testing community.

@@

From as early as the 1970s, a popular software development model was the waterfall model, where software was developed in a series of phases.  The waterfall development model is used because it is simple to implement correctly and is also considered time efficient because completing previous phases helps later phases become more efficient. The first phase in the standard waterfall model is the requirements phase where requirements are set by the customer or design company.  Next the design phase occurs in which the product is designed and then the product is built in the implementation phase.  The final phase is the verification phase where testing and debugging occurred \mycomment{wikipedia waterfall model}.  Due to the fact that these processes often put testing at the end of development, a certain type of testing method, the test-last method, became very popular to use with the waterfall development method.

Test-last testing is currently a popular testing method and is usually the first testing method that people tend to implement.  Test-last testing is the practice of writing tests after code has been written to check the functionality of the written code.  These tests are then used by the developer to fix their code until no errors are found by the tests.  /mycomment{this last sentence may no longer be needed} Test last testing allows the programmer to focus on writing tests based on areas of his code that he may be concerned about.

@@

In the late 1990s, some software developers started to criticize the sequence-oriented waterfall model, complaining that it was too brittle and inflexible to meet the demands of the standard customer.  In response to these critiques of the waterfall model, a new model for developing software emerged, the agile development model.  This new development model, based on the tenets of the agile manifesto \mycomment{cite manifesto here!}, promoted the idea that all actions of development should not occur within an ordered sequence of phases, but in a series of time-boxed iterations where in each iteration developers set requirements, design a product, make it, and test it.  The goal of each iteration is to produce a demonstrable working product to show a customer and to receive feedback on the product.  Some current development practices that are considered agile are Extreme Programming and Scrum.

@@

The most well known and most used test-first model is that of test driven development or TDD for short.  In its current use in the field, the phrase TDD is mostly used as a blanket term for any sort of development practice that uses test-first testing to drive code development. That being said it is important to note that although TDD has been turned into a vague term in this paper we will focus more on its original specific definition.  \mycomment{This paragraph will need to be reworked once the TDD section is reworked, it is currently out of date.}

\section{Research Data on Test-First vs Test-Last Testing}
In this section we will attempt to obtain useful comparisons between test-first and test-last testing by observing current research articles.  In order to accomplish this, this section starts with a data subsection that goes over three main studies that will be used to draw our conclusions.  Next, an analysis subsection summarizes the results of the studies in terms of our three main empirical measurements: code coverage, code correctness, and total development time.  Afterwards a discussion subsection talks about the potential issues with summarizing the data given in the research articles and then the conclusion subsection talks about potential useful comparisons we can draw from the data.

\subsection{Data}
\mycomment{Reworked the paragraphs on the Kollanus study and the Lemos study in an attempt to better represent data in my paper, also planning on writting paragraphs for two other studies.  Considering making this its own section due to its increase in size.}

This subsection contains summaries of important studies that will be used later in the paper to draw out important conclusions and comparisons between test-first and test-last testing.

The first research summarized research article  is a 2010 article by Kollanus ~\cite{Kollanus:2010}.  In this article, Kollanus, using standard literary review methods, found and reviewed forty different experiments in scientific journals, magazines, and conference proceedings that provided empirical evidence about the effectiveness of using TDD compared to test-last development methods.  In the review Kollanus focuses on three different measurements: external code quality, productivity, and internal code quality.  In the article, Kollanus describes external code quality as the measurement of the number of passed researcher acceptance tests in a test set when run against code written by participants.  Since this definition is equivalent to our definition for code-correctness this data can be used for code-correctness comparisons. In the study, Kollanus concluded that there was weak support for improved external code quality in TDD methods compared to test-last development methods.  This conclusion was based on consistent evidence found in case studies that reported better external code quality in TDD methods.  However, the support was considered weak because there were contradictory results found in some controlled experiments that recorded external code quality.  Productivity, as defined by Kollanus, is a measurement that measures how much code one writes in comparison to time spent.  Since this definition is similar to that of total development time's, this data can tentatively be used for total development time comparisons.  The research viewed by Kollanus heavily suggests that TDD methods are less productive and therefore take more time overall to write.  This conclusion was drawn from the fact that the majority of studies and experiments showed a decrease in productivity.  That being said multiple studies seem to contradict the majority.   Internal code quality describes a wide set of measurements that measure code quality from a development standpoint.  Although code coverage is one of these measurements and there is a mention that one article found increased code coverage in TDD methods, this sections conclusions will not be discussed as it does not fit with our paper's themes.

In 2012, Hellmann et. al conducted a summary on the current status of research of TDD.  Within Hellmann's background there is a mention of a summary study produced by Jeffries and Mitnik that found in general that TDD largely resulted in an increase in quality, but one study they identified showed instead that TDD resulted in a strong negative impact on quality.  Additionally while they showed that TDD could reduce the amount of effort required by up to 27\%, most studies found an increase in effort of up to 100\%.

In 2012, Lemos et. al ~\cite{Lemos:2012} conducted a study on third year computer science undergraduate students to see if test-first testing would significantly impact code coverage,  code correctness, and/or total development time in auxiliary functions (functions with 10-200 lines of code) compared to test-Last testing.   In this study 39 third year undergrad computer science students, knowledgeable in testing techniques and participants of two 100 minute long test first modules, were asked to complete coding challenges over two sessions.  In the first session, half the students implemented testing-first testing while the other half implemented test-last testing to help solve the problem.  In the second section the students were asked to switch roles and given a different coding challenge.  In order to reduce domain knowledge bias amongst the coding challenges three different challenge question types were handed out at each session.  Code coverage was measured by comparing the percentage of tested lines in participants code.  Total development time was measured by how long it took participants to finish their code.  Code correctness was measured by running acceptance tests against a participants development code and giving the code a score of 0(all test cases fail), .5(some test cases fail) or 1(all test cases pass).  Note this style of measuring code correctness has been used in previous studies and is call the \textit{Functional Test Set Success Level} scale.   The results of the study were analyzed for statistical significance using the Wilcoxon/Mann-Whitney non-parametric signed-rank paired test. For the results of code coverage, it was found that code coverage on average was 40\% higher when test-first testing was used.  On total development time, it was found that test-first code took 12\% longer to write then test-last code. Both of these results were statistically significant.  In terms of code correctness, the only difference found was that the test-last code had one more correct implimentation of code then the test-first code.  This result was not found to be statistically significant.

\mycomment{need to add two paragraphs for the studies by Janzen and Kettenun}

@@

\section{Test Driven Development}

\mycomment{currently this section is more of a stub then anything. This section should be fleshed out over the weekend.  Although I am not sure on formatting or precise content yet, many articles have noted that TDD takes longer and is harder to implement because it is unhelpful in determining what parts of the code/problem need to be tested and what is the best way of implementing those tests.} 
Test driven development, as mentioned earlier is currently the most widely used test first testing and design methodology.  Although in the previous cases we have been using TDD as a vague term in this section we will focus on the the original TDD methodology introduced in Extreme Programming.  In the original TDD methodology the developer repeats a step of processes to develop his code.  The process begins once all current tests pass or the new code has just been started.  When this occurs the programmer writes a new failing test that tests the simplest functionality the programmer wishes to add to their code.  Once the test has been written and the test fails, the programmer then writes the minimal amount of code to make the test pass.  After this step the programmer streamlines his solution with the other parts of his code.  This series of steps repeats until the code is complete.

~\mycomment{this next paragraph is an extremely rough draft that was written before this section became a section and currently makes little to no sense, comments are not needed for this paragraph}

Although TDD is a very popular style of test first testing, it should be noted that a perceived weakness to the TDD methodology exists.  The main issue many people have with test driven development is that it fails to suggest a style of testing and how to best define an independent functionality in ones code plan.  In the study ~\cite{Kettunen:2010} (cite others here as needed) people noted that TDD was difficult because they had a hard time deciding how to implement their tests effectively.  Another related issue noted in ~\cite{Kettunen:2010} was that some individuals found it hard to figure out what needed to specifically be tested thus some found they wrote an overabundance of tests causing the process to become inefficient and thus take more time.  Others found that focusing on the simplest functions first left serious potential problems to the end of development causing a serious time crunch problem for product release.  Due to these issues with defining functionality and other issues within TDD new methods for implementing test first testing are being designed and implemented in an attempt to solve these problems.

@@

\subsection{Acceptance Test Driven Development}
